---
title: "Mission Antarctica"
author: "Ellen Bledsoe"
date: "8/17/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Load the packages we will need:
```{r}
library(tidyverse)
library(maps)
library(patchwork)
library(cowplot)
```

## Tucson Airport Temperature Data

* Average monthly temperature (Celsius) data from the Tucson Airport 
* [Downloaded from NASA](https://data.giss.nasa.gov/cgi-bin/gistemp/stdata_show_v4.cgi?id=USW00023160&dt=1&ds=14) 

```{r}
tucson <- read_csv("../data/Tucson_Intl_Airport_WS.csv")
```

Let's take a look at the data.
```{r}
glimpse(tucson)
```
Note the 999.9s in the data. This is a common way of indicating that there is no data, or NA.
Let's tell R to convert 999.9 to NAs.

```{r}
tucson <- read_csv("../data/Tucson_Intl_Airport_WS.csv",
                   na = "999.9") %>% 
  rename(Summer = `J-J-A`, Year = YEAR)
glimpse(tucson)
```
That's better!

Now, let's plot the summer data. This is when the temps are hottest. It is also monsoon season, so the humidity is higher. This _could_ lead to "wet-bulb" temperatures of 35 C, which are extremely dangerous to humans.
```{r}
ggplot(data = tucson, aes(x = Year, y = Summer)) +
  geom_point() 
```

It's a good start, but let's make it better.
```{r}
ggplot(data = tucson, aes(x = Year, y = Summer)) +
  geom_point() +
  geom_line() +
  stat_smooth(method = 'lm') +
  labs(
    title = "Mean Summer Temp. at TUS",
    subtitle = "1949-2021",
    x = "Year",
    y = expression("Mean Summer Temp ("*degree*"C)")
  ) +
  theme_bw()

```

There's a clear upward trend in temperatures, which makes sense.

Now let's extrapolate. When does the average summer temperature become 35 C?

* NOTE: I'm not suggesting that this is the best method for prediction! This is simply an example of something you _can_ do.

```{r}
temp <- ggplot(data = tucson, aes(x = Year, y = Summer)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 35) + # 35 wet-bulb is too much
  xlim(1945, 2200) +
  stat_smooth(method = 'lm', fullrange = TRUE) +
  labs(
    title = "Mean Summer Temp. at TUS",
    subtitle = "1949-2021",
    x = "Year",
    y = expression("Mean Summer Temp ("*degree*"C)")
  ) +
  theme_bw()
temp
```

Looks to be around the year 2180.

Let's make a map with Tucson highlighted so people know what location we are talking about.
```{r}
# get data for map of US
us_states <- map_data("state") %>% 
  rename(Latitude = lat, Longitude = long)

# make a dataframe that has TUS location
tucson_airport <- data.frame(
  Latitude = c(32.1314),
  Longitude = c(-110.9553)
)

# plot map with TUS
map <- ggplot(us_states) +
  geom_polygon(aes(x = Longitude, y = Latitude, group = group), 
               fill = "white", color = "black") +
  coord_quickmap() + 
  geom_point(tucson_airport, mapping = aes(x = Longitude, y = Latitude), 
             color = "red", size = 5, shape = 18) +
  theme_classic() +
  theme(panel.border = element_rect(color = "black", fill = NA))
map
```

We can plot these next to each other
```{r}
plot <- map + temp
plot
```

Maybe better way to incorporate the map is as an insert?
```{r}
# the ggdraw() and draw_plot() functions are from the `cowplot` package
ggdraw() +
  draw_plot(temp) +
  draw_plot(map, height = 0.3, width = 0.3, x = 0.68, y = 0.125)
  
```

Other extrapolation analysis options: ARIMA
```{r}
library(forecast)

# time-series analysis
summer_ts <- ts(tucson$Summer)  # make temp data into a time series
model <- auto.arima(summer_ts)  # find the best ARIMA for the data
plot(forecast(model, h = 20))   # we can use the ARIMA to forecast 20 years out
plot(forecast(model, h = 200))  # also 200 years out
abline(h = 35)                  # add a line that indicates our cut-off
```






